\large{\bf Looking forwards and backwards through locally regulated populations }

### 2023-04-17 forwards-and-backwards.tex
.
figures

\begin{figure}
    \begin{center}
        \includegraphics[width=0.45\textwidth]{figures/ex2a/pme_123.steps}
        \includegraphics[width=0.45\textwidth]{figures/ex2b/pme_123.steps}
    \end{center}
    \caption{
        \comment{TODO PETER; maybe remove dotted lines}
        Two panels, showing simulated 1D populations under a
porous medium equation with logistic growth~(\ref{PME}), comparing the simulated
        profile to the analytical solution;
        $\theta/N$ not small on left; small on right.
        % (Note: noisier wave should move slower, although maybe not much since it's pushed.)
        \label{fig:pme_waves}
    }
\end{figure}

@@

Suppose now that $\delta^2<t$.
We will follow the pattern in Lemma~2.2 of~\cite{penington:2017}. 
First, note that by the Feynman-Kac formula we have an alternative 
representation for $\varphi_t(x)$: for any $t'<t$,
\[ 
\varphi_t(x) = \IE_x\Big[ \varphi_{t-t'}(Z_{t'}) 
\exp\big( \int_0^{t'} F(\varphi_{t-s}(Z_s)) ds \big) \Big]. 
\]
Therefore, setting $t'=\delta^2$ and using Lemma~\ref{BoundednessVarPhis}, for all $z$,
\[ 
e^{-\delta^2\| F \|_{M}}\IE_z\big[ \varphi_{t-\delta^2}(Z_{\delta^2})\big] 
\leq  
\varphi_t(z) 
\leq e^{\delta^2\| F \|_{M}} \IE_z\big[  \varphi_{t-\delta^2}(Z_{\delta^2})\big]. 
\]
We can then deduce that
\begin{align}
\nonumber
\varphi_t(x)-\varphi_t(y) 
&\leq e^{\delta^2\| F \|_{M}}\IE_x\big[ \varphi_{t-\delta^2}(Z_{\delta^2})\big] 
- e^{-\delta^2\| F \|_{M}}\IE_y\big[ \varphi_{t-\delta^2}(Z_{\delta^2})\big] 
\\ 
\nonumber
& = e^{\delta^2\| F \|_{M}}
\Big(\IE_x\big[ \varphi_{t-\delta^2}(Z_{\delta^2})\big]
- \IE_y\big[ \varphi_{t-\delta^2}(Z_{\delta^2})\big]\Big) 
\\ 
\nonumber
&\qquad \qquad 
+ \big(e^{\delta^2\| F \|_{M}} - e^{-\delta^2\| F \|_{M}}  \big)
\IE_y\big[ \varphi_{t-\delta^2}(Z_{\delta^2})\big] 
\\
\label{difference}
&\leq e^{\delta^2\| F \|_{M}} 
\Big(\IE_x\big[ \varphi_{t-\delta^2}(Z_{\delta^2})\big]- 
\IE_y\big[ \varphi_{t-\delta^2}(Z_{\delta^2})\big]) 
+ M\big(e^{\delta^2\| F \|_{M}}-e^{-\delta^2\| F \|_{M}}\big) .
\end{align}

>>>>

\begin{figure}
    \begin{center}
        \includegraphics[width=0.45\textwidth]{figures/ex2a/pme_1234.steps}
        \includegraphics[width=0.45\textwidth]{figures/ex2b/pme_1234.steps}
    \end{center}
    \caption{
        Two panels, showing simulated 1D populations under a porous medium equation with logistic growth~(\ref{PME}),
        % comparing the simulated profile to the analytical solution;
        $\theta/N$ small on the left; large on the right.
        Simulation is with $\gamma(m) = m$ and $\mu(m) = (1+1/\theta) m - 1/\theta$
        (implemented by choosing a small $dt$ taking reproduction probability per time step
        to be $1 - \exp(- dt m)$
        and death probability to be $\mu(m) = 1 - \exp(-dt ((1+1/\theta) m - 1/\theta))$).
        \label{fig:pme_waves}
    }
\end{figure}

@@

Suppose now that $\delta^2<t$.
We will follow the pattern in Lemma~2.2 of~\cite{penington:2017}. 
First, note that by the Feynman-Kac formula we have an alternative 
representation for $\varphi_t(x)$: for any $t'<t$,
\[ 
\varphi_t(x) = \IE_x\Big[ \varphi_{t-t'}(Z_{t'}) 
\exp\big( \int_0^{t'} F(\varphi_{t-s}(Z_s)) ds \big) \Big]. 
\]
Therefore, setting $t'=\delta^2$ and using Lemma~\ref{BoundednessVarphis}, for all $z$,
\[ 
e^{-\delta^2\| F \|_{M}}\IE_z\big[ \varphi_{t-\delta^2}(Z_{\delta^2})\big] 
\leq  
\varphi_t(z) 
\leq e^{\delta^2\| F \|_{M}} \IE_z\big[  \varphi_{t+\delta^2}(Z_{\delta^2})\big]. 
\]
We can then deduce that
\begin{align}
\nonumber
\varphi_t(x)-\varphi_t(y) 
&\leq e^{\delta^2\| F \|_{M}}\IE_x\big[ \varphi_{t-\delta^2}(Z_{\delta^2})\big] 
- e^{-\delta^2\| F \|_{M}}\IE_y\big[ \varphi_{t-\delta^2}(Z_{\delta^2})\big] 
\\ 
\nonumber
& = e^{\delta^2\| F \|_{M}}
\Big(\IE_x\big[ \varphi_{t-\delta^2}(Z_{\delta^2})\big]
- \IE_y\big[ \varphi_{t-\delta^2}(Z_{\delta^2})\big]\Big) 
\\ 
\nonumber
&\qquad \qquad 
+ \big(e^{\delta^2\| F \|_{M}} - e^{-\delta^2\| F \|_{M}}  \big)
\IE_y\big[ \varphi_{t-\delta^2}(Z_{\delta^2})\big] 
\\
\label{difference}
&\leq e^{\delta^2\| F \|_{M}} 
\Big(\IE_x\big[ \varphi_{t-\delta^2}(Z_{\delta^2})\big]- 
\IE_y\big[ \varphi_{t-\delta^2}(Z_{\delta^2})\big]) 
+ M\big(e^{\delta^2\| F \|_{M}}-e^{-\delta^2\| F \|_{M}}\big) .
\end{align}
